### 2015-04-01 dAlg/mrm.tex
MRM template for paper sketching

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[nocopyrightspace]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{xcolor}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{doubleequals}
\usepackage{comment}
\usepackage{url}

%% for comments
\input{Preamble}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\def\commentbegin{\quad\{\ }
\def\commentend{\}}

\newcommand\SM[1]{\textcolor{red!70!yellow}{[SM: #1]}}
\newcommand\BO[1]{\textcolor{green!70!black}{[BO: #1]}}
\newcommand\SH[1]{\textcolor{blue!70!gray}{[SH: #1]}}

\newcommand\name{{\bf DSL}\xspace}
\newcommand\delete[1]{}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Composable Algebra with Dependencies}
%\subtitle{A List-of-Functors Approach to Two-Level Types}

\authorinfo{Bruno C. d. S. Oliveira}
           {University of Hong Kong}
           {bruno@cs.hku.hk}
\authorinfo{Shin-Cheng Mu}
           {Academia Sinica}
           {scm@iis.sinica.edu.tw}
\authorinfo{Shu-Hung You}
           {National Taiwan University}
           {suhorngcsie@gmail.com}
\maketitle

\begin{abstract}

\end{abstract}

%\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
%\terms
%term1, term2

%\keywords
%keyword1, keyword2

\input{sections/Introduction}

\input{sections/Overview}

\input{sections/ComposableAlg}

\input{sections/DependentAlg}

\input{sections/Extensibility}

\section{Other representations of algebra}
\label{sec: other-alg}

\input{sections/TypeClassAlg}

\input{sections/RecordAlg}

\input{sections/RelatedWork}

\input{sections/Conclusion}

\begin{comment}
\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.
\end{comment}

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\bibliography{mrm}

\end{document}

%                       Revision History
%                       -------- -------
%  Date         Person  Ver.    Change
%  ----         ------  ----    ------

%  2013.06.29   TU      0.1--4  comments on permission/copyright notices