### 2015-04-18 dAlg/mrm.tex
Wrap up

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[nocopyrightspace]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{xcolor}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{doubleequals}
\usepackage{comment}
\usepackage{url}
\usepackage{graphicx}
\usepackage[export]{adjustbox}
\graphicspath{ {images/} }

%% for comments
\input{Preamble}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\def\commentbegin{\quad\{\ }
\def\commentend{\}}

\newcommand\SM[1]{\textcolor{red!70!yellow}{[SM: #1]}}
\newcommand\BO[1]{\textcolor{green!70!black}{[BO: #1]}}
\newcommand\SH[1]{\textcolor{blue!70!gray}{[SH: #1]}}

\newcommand\name{{\bf DSL}\xspace}
\newcommand\delete[1]{}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{20yy} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Composable Algebra with Dependencies}
%\subtitle{A List-of-Functors Approach to Two-Level Types}

\authorinfo{Draft}
           {}
           {}

\maketitle

\begin{abstract}

This paper presents composable algebra with dependencies: a new approach to compose 
algebras together to allow dependent interpretations with {\em fold}. 
For a given datatype represented by two-level-types\cite{sheard04}, if one of its 
interpretations depends on a second one, we compose the algebra correponding 
to the first interpretation with the one corresponding to the second, and use the 
composed algebra with {\em fold} to evaluate the expression. We present our approach 
on top of the domain specific language for parallel prefix circuits.

\end{abstract}

%\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore, 
% you may leave them out
%\terms
%term1, term2

%\keywords
%keyword1, keyword2

\input{sections/Introduction}

\input{sections/Scans}

\input{sections/FAlg}

\input{sections/Overview}

\input{sections/ComposableAlg}

\input{sections/DependentAlg}

\input{sections/Extensibility}

\section{Other representations of algebra}
\label{sec: other-alg}

Apart from {\em two-level-types}, there are other ways to represent and evaluate
the DSL for parallel prefix circuits. In this section, we will show two other 
representations, and how they can be used to allow dependent interpretations.

\input{sections/TypeClassAlg}

\input{sections/RecordAlg}

\input{sections/RelatedWork}

\input{sections/Conclusion}

\begin{comment}
\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.
\end{comment}

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

%\bibliography{mrm}

\begin{thebibliography}{}
\softraggedright

\bibitem{sheard04}
  Sheard, Tim and Pasalic, Emir.
  Two-level types and parameterized modules.
  \emph{Journal of Functional Programming}.
  14(5):547-587, September 2014.

\bibitem{gibbons14}
  Gibbons, Jeremy and Wu, Nicolas.
  Folding Domain-Specific Languages: Deep and Shallow Embeddings.
  \emph{International Conference on Functional Programming}.
  2014.

\bibitem{fokkinga90}
  Fokkinga, Maarten M.
  Tupling and mutumorphisms.
  1990.

\bibitem{oliveira15}
  Oliveira, B.C.d.S., Mu, Shin-Cheng and You, Shu-Hung.
  Modular Reifiable Matching: A List-of-Functors Approach to Two-Level Types.
  February 2015.

\bibitem{bahr15}
  Bahr, Patrick and Axelsson, Emil.
  Generalising Tree Traversals to DAGs.
  \emph{Proceedings of the 2015 Workshop on Partial Evaluation and Program Manipulation},
  pages 27-38.
  2015

\end{thebibliography}

\end{document}

%                       Revision History
%                       -------- -------
%  Date         Person  Ver.    Change
%  ----         ------  ----    ------

%  2013.06.29   TU      0.1--4  comments on permission/copyright notices